---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  eval = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pgnapes

<!-- badges: start -->
<!-- badges: end -->

The goal of {pgnapes} is to:

* Establish R-connection to PGNAPES Oracle database
* Provide convenient functions to PGNAPES tables and views
* R-code examples for data analysis and graphical presentations

## Installation

You can install the development version of pgnapes from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
install.packages("devtools")
remotes::install_github("einarhjorleifsson/pgnapes")
```

The backbone of {pgnapes} is {dplyr} which allows for generation of simple or complex sequence of codes that are automatically translated to SQL via the {dbplyr}-package. Since the PGNAPES database is Oracle 11g one can not use {dbplyr} versions >2.0. Hence we need to install an older version:

```{r, eval = FALSE}
remotes::install_github("tidyverse/dbplyr@v1.4.4", force = TRUE)
```

**Side effects**: Installing an earlier version of {dbplyr} means that one can not load the {tidyverse} package only its child packages (dplyr, ggplot2, tidyr, tibble, ….). Reinstalling the latest version of {dbplyr} is done by `install.packages("dbplyr")`.


## Connection

All functions in the {pgnapes} package have a prefix `pgn_`. The core function in {pgnapes} is the `pgn_connect` that provides a connection to the PGNAPES database. The arguments of the function are *username*, *password* and *use_odbc*. The two first arguments are self-expanatory. The arguement *use_odbc* (default TRUE) provides a control if connection is made via {odbc} or {ROracle}. The latter is generally faster but the {ROracle}-package setup is more difficult (see furhter notes below)

**At the start of an R-sesssion load some packages**:

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(maps)
library(pgnapes)
```

**Connect to the database**:

The first thing to do is to establish a connection to the PGNAPES-database by:

```{r, eval = FALSE}
con <- pgn_connect(username = "your_user_name", password = "your_password")
```

```{r, echo = FALSE}
load("leyndo.RData")
con <- pgn_connect(my_user_name, my_password, use_odbc = TRUE)
```

You should now get something like the following:

```{r}
con
```

## The PGNAPES tables

The functions below are a convenient wrapper to access all the tables in pgnapes:

```{r, eval = FALSE}
pgn_acoustic(con)
pgn_acousticvalues(con)
pgn_biology(con)
pgn_catch(con)
pgn_countries(con)
pgn_hydrography(con)
pgn_icessquares(con)
pgn_icessquares_big(con)
pgn_igoss(con)
pgn_plankton(con)
pgn_logbook(con)
pgn_species(con)
pgn_stationtypes(con)
pgn_survey(con)
pgn_vessels(con)
```

## Examples

### Get and "standardize" mackerel catch

```{r}
d <-
  pgn_logbook(con) %>%
  filter(year >= 2013,
         month %in% 7:8,
         sttype %in% c("PTRAWL", "TRAWL", "Multpelt", "PTrawl_Straight",
                       "DeepTrawl", "Multpelt 832")) %>%
  arrange(cruise, year, month, day, hour, min) %>%
  left_join(pgn_catch(con) %>% 
              filter(species %in% "MAC")) %>%
  collect(n = Inf) %>% 
  mutate(catch = catch / 1000 / (towtime / 60),
         catch = replace_na(catch, 0))
glimpse(d)
```

**Map catch**:

```{r maccatch, fig.height = 11}
xlim <- range(d$lon)
ylim <- range(d$lat)
m <- map_data("world", xlim = xlim, ylim = ylim)
d %>% 
  ggplot(aes(lon, lat)) +
  theme_bw() +
  geom_polygon(data = m, aes(long, lat, group = group), fill = "grey") +
  geom_path(aes(group = cruise), colour = "grey", lwd = 0.25, alpha = 0.5) +
  geom_point(data = d %>% filter(catch > 0),
             aes(size = catch), colour = "red", alpha = 0.35) +
  scale_size_area(max_size = 5) +
  geom_point(size = 0.01, alpha = 0.2) +
  facet_wrap(~ year, ncol = 2) +
  #coord_quickmap(xlim = xlim, ylim = ylim) +
  coord_map(projection = "azequalarea",
            xlim = xlim, ylim = ylim) +
  scale_x_continuous(NULL, NULL) +
  scale_y_continuous(NULL, NULL)
```


## Intalling ROracle

Your IT-department may be of help here, details on intallments are [here](http://cran.hafro.is/web/packages/ROracle/INSTALL). Information on the installation of the needed Oracle Instant Client are [here](https://www.oracle.com/database/technologies/instant-client.html).
